---
title: "PA1_Templets"
author: "baki"
date: "6/2/2019"
output: html_document
---

---
title: "ReproducibleResearchProject1v2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reproducible Research Course Project 1 (Week 2)

## Loading and preprocessing the data
```{r Question 1}
fitnessdata <- read.csv("activity.csv", header = TRUE)
head(fitnessdata)
```

## What is mean total number of steps taken per day?
### Calculate the total number of steps taken per day
### Make a histogram of the total number of steps taken each day
```{r Calculate the total number of steps taken per day; Make a histogram of the total number of steps taken each day}
stepsbyday <- aggregate(steps ~ date, fitnessdata, FUN = sum, na.rm=TRUE)
hist(stepsbyday$steps, xlab = "Total daily Steps",main="Histogram of Total Steps by day", breaks = 20)
head(stepsbyday)
```

### Calculate and report the mean and median of the total number of steps taken per day
```{r Calculate and report the mean and median of the total number of steps taken per day}
print(mean(stepsbyday$steps))
print(median(stepsbyday$steps))
```

## What is the average daily activity pattern?
```{r avg active daily pattern}
#Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
library(ggplot2)
stepsbyinterval <- aggregate(steps ~ interval, fitnessdata, FUN = mean, na.rm=TRUE)
ggplot(stepsbyinterval, aes(x=interval, y=steps))+ geom_line()
```

```{r max interval}
#Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
stepsbyinterval[which(stepsbyinterval$steps== max(stepsbyinterval$steps)),]
```

## Imputing missing values
```{r  hist with NA mean filled}
fitnessdata_filled <- fitnessdata
nrow(fitnessdata_filled[is.na(fitnessdata_filled$steps),])
fitnessdata_narm <- fitnessdata_filled[!is.na(fitnessdata_filled$steps),]
missingData <- is.na(fitnessdata_filled$steps)
meanVals <- tapply(fitnessdata_narm$steps, fitnessdata_narm$interval, mean, na.rm=TRUE, simplify=TRUE)
fitnessdata_filled$steps[missingData] <- meanVals[as.character(fitnessdata_filled$interval[missingData])]
head(fitnessdata_filled)
write.csv(fitnessdata_filled, file = "activity_filled.csv")
stepsbyday_filled <- aggregate(steps ~ date, fitnessdata_filled, FUN = sum, na.rm=TRUE)
hist(stepsbyday_filled$steps, xlab = "Total daily Steps",main="Histogram of Total Steps by day (NA Mean Filled)", breaks = 20)
```

## Are there differences in activity patterns between weekdays and weekends?
#### 1 of 3
```{r weekday vs weekend 1 of 3}
# Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day. 
mean(stepsbyday$steps)
mean(stepsbyday_filled$steps)
median(stepsbyday$steps)
median(stepsbyday_filled$steps)
```

#### 2 of 3
```{r  weekday vs weekend 2 of 3}
#head(as.Date(stepsbyday_filled$date))
fitnessdata_filled$weekday <- weekdays(as.Date(fitnessdata_filled$date))
fitnessdata_filled$weekend <- ifelse (fitnessdata_filled$weekday == "Saturday" | fitnessdata_filled$weekday == "Sunday", "Weekend", "Weekday")
head(fitnessdata_filled,7)
```

#### 3 of 3
```{r weekday vs weekend 3 of 3}
weekendagg <- aggregate(steps ~ weekend + interval, fitnessdata_filled, FUN = mean)
names(weekendagg) <- c("weekend", "interval", "steps")
ggplot(weekendagg, aes(x = interval, y=steps, color=weekend)) + geom_line() + facet_grid(weekend ~ .)
```
# End